### 2018-03-17 paper/4-build.tex
Minor tweaks to S4

When using a dirty bit, it is necessary to check all the dependencies of a key. For applicative build systems that list is simple to obtain, but for monadic build systems there is no general way to get all dependencies. Instead \Excel computes a \emph{static approximation} of the dependencies. For applicative rules that approximation is correct. For functions such as \cmd{IF} it marks the cell dirty if \emph{any} potential dependency has changed, even on the untaken \emph{if} branch. For functions such as \cmd{INDIRECT} whose dependencies cannot be guessed, it conservatively assumes the dependencies have always changed.

@@

We assume that \hs{Hash v} is a small constant size, constructed from hashing the
underlying \hs{v} rather than storing it directly. Checking a trace requires
ensuring all the dependencies are up to date (using whatever ordering strategy
as per \S\ref{sec-dependency-orderings}), then comparing if the dependencies are
same as the current value and that the result is the same.

A verifying trace build system needs to maintain a set of traces. After computing a
fresh value we add its \hs{Trace k v} to the set. Therefore the information stored
by a build system that verifies traces can be stored as a list or set - we chose \hs{[Trace k v]} for simplicity. In practice, different build systems can optimise the data structures used by traces for their specific use cases, which we discuss in \S\ref{sec-smart-traces}.

@@

\begin{table}[h]
\smaller
\centering
\begin{tabular}{l||c|c|c}
\hline
Property           & Topological \S\ref{sec-topological} & Reordering \S\ref{sec-reordering} & Recursive \S\ref{sec-recursive}    \\\hline
\hline
Dirty bit \S\ref{sec-dirty-bit}                    & \Make    & \Excel                   & Approximate \Shake  \\\hline
Verifying trace \S\ref{sec-verifying-traces}       & \Ninja   & Traced \Excel            & \Shake                      \\\hline
Constructive trace \S\ref{sec-constructive-traces} & \Bazel   & Cloud \Excel             & Cloud \Shake                \\\hline
\end{tabular}
\vspace{0.5mm}
\caption{Build systems \`a la Carte.\label{tab-build-systems}}
\end{table}

With the information in this section we can build a table comparing the dependency order strategy with the out-of-date keys strategy, providing 9 possible build systems, 5 of which are actually inhabited by existing build systems (we discuss \Ninja \cite{ninja} in \S\ref{sec-related-build}). Of the remaining 4 spots, we believe neither Traced or Cloud \Excel make sense - the \Excel approach of reordering combined with static approximations reduces the memory usage significantly. However, as soon as you are paying the cost of storing traces, that benefit is gone. The advantage of Approximate \Shake over \Shake would be that it could avoid storing traces and having a separate information database, but that advantage is minor compared to the technical restrictions and approximations it would provide, so we consider it unlikely to be built. The Cloud \Shake system seems like an interesting and important point in the design space, which we explore further in \S\ref{sec-implementation-cloud-shake}.

>>>>

When using a dirty bit, it is necessary to check all the dependencies of a key.
For applicative build systems that list is easy to obtain, but for monadic
build systems there is no general way to get all dependencies. Instead \Excel
computes a \emph{static approximation} of the dependencies. For applicative
tasks that approximation is correct. For functions such as \cmd{IF} it marks the
cell dirty if \emph{any} potential dependency has changed, even on the untaken
\emph{if} branch. For functions such as \cmd{INDIRECT} whose dependencies cannot
be guessed, it conservatively assumes the dependencies have always changed.

@@

We assume that \hs{Hash}~\hs{v} is a small constant size, constructed from hashing the
underlying \hs{v} rather than storing it directly. Checking a trace requires
ensuring all the dependencies are up to date (using whatever ordering strategy
as per \S\ref{sec-dependency-orderings}), then comparing if the dependencies are
same as the current value and that the result is the same.

A build system that uses verifying traces needs to persistently maintain a set
of traces. After computing a fresh value we add its \hs{Trace}~\hs{k}~\hs{v} to
the set. Therefore the information stored by a build system that verifies
traces can be stored as a list or set~--~we chose \hs{[Trace}~\hs{k}~\hs{v]}
for simplicity. In practice, different build systems can optimise the data
structures used by traces for their specific use cases, which we discuss
in~\S\ref{sec-smart-traces}.

@@

\begin{table}[h]
\smaller
\centering
\begin{tabular}{l||c|c|c}
\hline
Property           & Topological\hspace{2mm}\S\ref{sec-topological} & Reordering\hspace{2mm}\S\ref{sec-reordering} & Recursive\hspace{2mm}\S\ref{sec-recursive}    \\\hline
\hline
Dirty bit\hfill\S\ref{sec-dirty-bit}                    & \Make    & \Excel                   & Approximate \Shake*  \\\hline
Verifying trace\hfill\S\ref{sec-verifying-traces}       & \Ninja   & Traced \Excel*            & \Shake                      \\\hline
Constructive trace\hspace{2mm}\hfill\S\ref{sec-constructive-traces} & \Bazel   & Cloud \Excel*             & Cloud \Shake*                \\\hline
\end{tabular}
\vspace{2mm}
\caption{Build systems \`a la carte. Systems marked * are hypothetical systems that do not currently exist.\label{tab-build-systems}}
\vspace{-2mm}
\end{table}

With the information in this section we can build a table comparing the
dependency order strategy with the out-of-date keys strategy, providing 9~possible
build systems, 5~of which are actually inhabited by existing build systems
(we discuss \Ninja \cite{ninja} in \S\ref{sec-related-build}). Of the
remaining 4~spots, we believe neither Traced or Cloud \Excel make sense~--~the
\Excel approach of reordering combined with static approximations reduces
the memory usage significantly. However, as soon as you are paying the cost of
storing traces, that benefit is gone. The advantage of Approximate \Shake over
\Shake would be that it could avoid storing traces and having a separate
information database, but that advantage is minor compared to the technical
restrictions and approximations it would provide, so we consider it unlikely
to be built. The Cloud \Shake system is an interesting and important point in
the design space, which we explore further in~\S\ref{sec-implementation-cloud-shake}.
